---
title: "Question 2 - Analysis"
output: html_notebook
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# QUESTION 2 (using “field data.xlsx”)

The first sheet of the data provided shows contact history for organizers working on a voter registration
campaign. The second sheet shows the weekly goal by region for number of phone attempts. Using this,
create a report for the campaign’s field director that summarizes each region’s progress towards hitting
their weekly phone attempt goals and any recommendations you may have for them. Note: you may want
to look at the progress of individual organizers. The report should be no more than 3 pages and may
include charts/visualizations of your choosing.

# Load packages
```{r}
# knitr: This package is used to embed R code in LaTeX, HTML, Markdown, and other formats for dynamic report generation.
library(knitr)

# tidyverse: This package is a collection of R packages designed for data science. It includes dplyr, tidyr, ggplot2, readr, purrr, tibble, stringr, and forcats.
library(tidyverse)

# readxl: This package makes it easy to read Excel files in R.
library(readxl)

# ggplot2: This package is part of the tidyverse and is used for data visualization. It implements the grammar of graphics, a coherent system for describing and building graphs.
library(ggplot2)

# extrafont: This package allows one to use additional fonts in plots created with ggplot2 and other R graphics functions, enabling customization of text appearance.
library(extrafont)
```

## Summarize the number of phone attempts made by each region and week
```{r}
# Read in the data
field_data <- read_xlsx("field data.xlsx", sheet = "Data") # "Data" sheet 
phone_goals <- read_xlsx("field data.xlsx", sheet = "Phone Goals") # "Phone Goals" sheet

# Filter for phone attempts
phone_attempts <- field_data %>%
  filter(ContactType == "Phone")

# Summarize the number of phone attempts by region and week
phone_attempts_summary <- phone_attempts %>%
  group_by(Region, Week, ContactType) %>%
  summarize(Attempts = n(), .groups = 'drop')

# Join the summary with phone goals
summary_with_goals <- phone_attempts_summary %>%
  left_join(phone_goals, by = c("Region", "Week", "ContactType"))

# Calculate progress towards goals
summary_with_goals <- summary_with_goals %>%
  mutate(Progress = Attempts / `Attempt Goal` * 100)

# Convert Week and Region to a factor
summary_with_goals$Week <- as.factor(summary_with_goals$Week)
summary_with_goals$Region <- as.factor(summary_with_goals$Region)

# View the summarized data
print(summary_with_goals)

# Save CSV of summary for tables
write.csv(summary_with_goals, "Region Summary.csv")
```
### Create visualizations to show each region's overall progress
```{r}
# Define a color palette with distinct colors for each region
region_colors <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b")

overall_progress <- ggplot(summary_with_goals, aes(x = Week, y = Progress, color = Region, group = Region)) +
  geom_line(size = 1) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "red") +
  geom_point(size = 2.5) +
  scale_color_manual(values = region_colors) +  # Apply the custom color palette
  labs(title = "Overall Progress Towards Weekly Phone Attempt Goals by Region",
       x = "Week",
       y = "Progress (%)",
       color = "Region") +
  theme_minimal() +
  theme(strip.text = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold"),
        panel.grid = element_blank(),
        panel.background = element_blank(),    # Remove background
        plot.background = element_blank(),      # Remove plot background
        legend.position = "right",
        text = element_text(family = "Trebuchet MS"))

print(overall_progress)

ggsave("overall_progress.png", overall_progress, width = 10, height = 6, dpi = 300)
```
### Create visualizations to show each region's progress
```{r}
region_progress <- ggplot(summary_with_goals, aes(x = Week, y = Progress)) +
  geom_bar(stat = "identity", aes(fill = ifelse(Progress >= 100, "Meeting Goal", "Not Meeting Goal"))) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "red") +
  facet_wrap(~ Region) +
  scale_fill_manual(values = c("Not Meeting Goal" = "red", "Meeting Goal" = "gray")) +
  labs(title = "Progress Towards Weekly Phone Attempt Goals by Region",
       x = "Week",
       y = "Progress (%)",
       fill = "Goal Status") +
  theme_minimal() +
  theme(strip.text = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold"),
        panel.grid = element_blank(),          # Remove grid lines
        panel.background = element_blank(),    # Remove background
        plot.background = element_blank(),      # Remove plot background
        text = element_text(family = "Trebuchet MS")
        )

print(region_progress)

# Save the modified plot
ggsave("region_progress.png", region_progress, width = 12, height = 8, dpi = 300)
```
## Summarize the number of phone attempts made by each organizer and week
```{r}
# Summarize the number of phone attempts by organizer, region, and week
organizer_summary <- phone_attempts %>%
  group_by(Organizer, Region, Week, ContactType) %>%
  summarize(Attempts = n(), .groups = 'drop')

# Join the summary with phone goals
organizer_summary <- organizer_summary %>%
  left_join(phone_goals, by = c("Region", "Week", "ContactType"))

# Calculate progress towards goals for each organizer
organizer_summary <- organizer_summary %>%
  mutate(Progress = Attempts / `Attempt Goal` * 100)

# Convert Week and Region to factors
organizer_summary$Week <- as.factor(organizer_summary$Week)
organizer_summary$Region <- as.factor(organizer_summary$Region)

# View the summarized data
print(organizer_summary)

# Save CSV of summary for tables
write.csv(organizer_summary, "Organizer Summary.csv")
```
### Create heatmap of towards goals by organizers
```{r}
organizer_progress <- ggplot(organizer_summary, aes(x = Week, y = reorder(Organizer, desc(Organizer)), fill = Progress)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "red", high = "green", na.value = "grey50") +
  labs(title = "Progress Towards Weekly Phone Attempt Goals by Organizer",
       x = "Week",
       y = "Organizer ID",
       fill = "Progress (%)") +
  theme_minimal() +
  theme(text = element_text(family = "Trebuchet MS"),
        strip.text = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold"),
        panel.grid = element_blank(),          # Remove grid lines
        panel.background = element_blank(),    # Remove background
        plot.background = element_blank(),      # Remove plot background
  )

print(organizer_progress)

ggsave("organizer_progress.png", organizer_progress, width = 12, height = 6, dpi = 300)
```
